### 2023-10-20 tabs/top10cat_description.tex
Make table for traffic to top ten categories

                                  Business & 720,930 & 13.0 &   720,930 & 13.0 & (decipherinc.com, samplicio.us, privatelink.de) & 4.8 & 10.0 & 0.0 & 0.5 & 1.8 & 5.0 & 11.7 & 18.7 & 140.5 \\
                Search Engines and Portals & 670,710 & 12.1 & 1,391,640 & 25.0 &           (google.com, google.co.uk, yahoo.com) & 4.1 &  8.9 & 0.0 & 0.2 & 1.5 & 4.9 & 10.2 & 15.4 & 139.9 \\
                Chat and Instant Messaging & 613,394 & 11.0 & 2,005,034 & 36.0 &               (google.com, yahoo.com, live.com) & 9.0 & 25.8 & 0.0 & 0.2 & 2.7 & 9.7 & 21.3 & 31.1 & 492.3 \\
               Business, Social Networking & 443,281 &  8.0 & 2,448,315 & 44.0 &        (facebook.com, facebook.co, soocial.com) & 6.4 & 18.6 & 0.0 & 0.1 & 0.6 & 4.5 & 15.4 & 33.0 & 245.5 \\
                                  Shopping & 328,300 &  5.9 & 2,776,615 & 49.9 &             (amazon.com, ebay.com, walmart.com) & 3.6 &  7.3 & 0.0 & 0.2 & 1.2 & 3.8 &  9.0 & 15.9 &  79.0 \\
          Business, Information Technology & 284,428 &  5.1 & 3,061,043 & 55.0 &       (clarity.ms, sentry.io, inboxdollars.com) & 1.8 &  4.9 & 0.0 & 0.1 & 0.5 & 1.8 &  4.1 &  7.0 &  87.8 \\
            Entertainment, Streaming Media & 261,165 &  4.7 & 3,322,208 & 59.7 &            (youtube.com, hulu.com, netflix.com) & 7.2 & 32.7 & 0.0 & 0.1 & 0.3 & 3.0 & 14.8 & 27.8 & 617.5 \\
News and Media, Search Engines and Portals & 231,157 &  4.2 & 3,553,365 & 63.9 &                             (bing.com, att.net) & 3.2 &  6.3 & 0.0 & 0.0 & 0.7 & 3.8 &  9.0 & 16.0 &  61.8 \\
                       Business, Education & 181,106 &  3.3 & 3,734,471 & 67.1 &           (yougov.com, google.com, prolific.co) & 1.8 &  4.9 & 0.0 & 0.2 & 0.7 & 1.8 &  3.8 &  6.5 & 104.5 \\
                        Business, Shopping & 135,431 &  2.4 & 3,869,902 & 69.6 &        (amazon.com, rakuten.com, instacart.com) & 1.5 &  3.2 & 0.0 & 0.1 & 0.4 & 1.3 &  3.5 &  5.9 &  32.2