### 2024-04-15 ms/lit_review.tex

### Violence in videos

"Of the 304 scenes analyzed, 88.2\% contained physical aggression, principally spanking, gagging, and slapping, while 48.7\% of scenes contained verbal aggression, primarily name-calling. Perpetrators of aggression were usually male, whereas targets of aggression were overwhelmingly female." \citep{bridges2010aggression}
    - Data:  The population of titles for this research was drawn from a compilation of 250 best-selling and 250 most rented video lists published monthly by AVN. The researchers selected the top 30 videos appearing on each list from December 2004 to June 2005. After deleting duplications, the population consisted of 275 titles (AVN, 2005). Fifty titles were randomly selected from this list to comprise the sample, yielding 304 scenes ...

### Correlates of Porn

"Nonetheless, demand for several online services does not appear to be sensitive to income, including chat, credit history, social media, and pornography." \citep{boik2016empirical}